---
title: 'Origin-destination explorer: a Shiny app'
author: "Yuan Liao"
date: "10/28/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(sf)
library(stplanr)
library(tmap)
```

## Transport, the big source of carbon emissions
- "After 2030, transport is likely to become the sector with the highest emissions in the 2-degree scenario." -- Climate Change 2013 by IPCC

- Increasing attention on a mode shift from car to active travelling such as biking among other low-carbon modes in cities.

## How do the travel demand by biking and other modes look like?
[Interactive ODM explorer](https://yuanliao.shinyapps.io/InteractiveODM/) is featured with:

- Origin-destination matrix broken down by mode and time

- Interactive visualization of OD network

- Using the travel survey data collected from the Netherlands (North Holland area) which is famous for its popularity of biking

## Interactive ODM explorer - Demo 1

-> Below is an example of the output: 

```{r demo, echo = TRUE, warning=FALSE, message=FALSE}
df <- read.csv('InteractiveODM/data/odms_by_mode.csv')
colnames(df) <- c("ozone", "dzone", "Non-bike", "E-bike", "Bike", "Total", 'Hour')
zones <- st_transform(st_read('InteractiveODM/data/zones_subset.shp', quiet = TRUE), crs = 4326)
df <- df %>% filter(ozone %in% zones$zone, dzone %in% zones$zone)

input <- list()
input$mode <- "Non-bike"
input$hour <- 24

df_a <- df %>%
  select(c('ozone', 'dzone', !!input$mode, 'Total', 'Hour')) %>%
  filter(ozone != dzone,
         Hour == input$hour) %>%
  mutate(trip = get(input$mode) / sum(get(input$mode)) * 100) %>%
  filter(trip > 0) %>%
  mutate(share = get(input$mode) / Total * 100)

l <- od2line(flow = df_a, zones = zones)

l <- l %>%
  arrange(trip)

tm_shape(zones) +
  tm_fill(col="grey35") +
  tm_borders("white", alpha=.2) + 
tm_shape(l) +
  tm_lines(
    palette = "plasma", 
    trans = "log10", style="cont",
    lwd = "trip",
    scale = 9,
    title.lwd = 0.5,
    alpha = 0.5,
    col = "share",
    title = "Mode share (%)",
    legend.lwd.show = FALSE
  ) +
  tm_scale_bar() +
  tm_layout(
    frame = FALSE,
    legend.bg.alpha = 0.5,
    legend.bg.color = "white"
  )
```

## Interactive ODM explorer - Demo 2
-> Below is another example of the output: the Bike trips at 8 am

```{r demo2, echo = FALSE, warning=FALSE, message=FALSE}
input <- list()
input$mode <- "Bike"
input$hour <- 8

df_a <- df %>%
  select(c('ozone', 'dzone', !!input$mode, 'Total', 'Hour')) %>%
  filter(ozone != dzone,
         Hour == input$hour) %>%
  mutate(trip = get(input$mode) / sum(get(input$mode)) * 100) %>%
  filter(trip > 0) %>%
  mutate(share = get(input$mode) / Total * 100)

l <- od2line(flow = df_a, zones = zones)

l <- l %>%
  arrange(trip)

tm_shape(zones) +
  tm_fill(col="grey35") +
  tm_borders("white", alpha=.2) + 
tm_shape(l) +
  tm_lines(
    palette = "plasma", 
    trans = "log10", style="cont",
    lwd = "trip",
    scale = 9,
    title.lwd = 0.5,
    alpha = 0.5,
    col = "share",
    title = "Mode share (%)",
    legend.lwd.show = FALSE
  ) +
  tm_scale_bar() +
  tm_layout(
    frame = FALSE,
    legend.bg.alpha = 0.5,
    legend.bg.color = "white"
  )
```

